---
title: "Colchicine to Suppress Pro-Tumorigenic Inflammation in Advanced Solid Tumors and Resected High-Risk Urothelial Carcinoma: A Pilot Trial"
subtitle: "Clinical Trial Results (The Oncologist CTR format)"
author: "Alexander B. Karol; Che-Kai Tsao, MD; Himanshu Joshi, PhD; Deborah Y. Doroshow, MD, PhD; Matthew D. Galsky, MD; Thomas Marron, MD, PhD"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    theme: flatly
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 7, fig.height = 5)
set.seed(123)

suppressPackageStartupMessages({
  library(here)
  library(readr)
  library(dplyr)
  library(tidyr)
  library(stringr)
  library(lubridate)
  library(ggplot2)
  library(knitr)
  library(kableExtra)
  library(survival)
  library(survminer)
})

# ---- Load REDCap export and label helpers ----
source(here::here("scripts", "protocol2101175_redcap_helpers.R"))
data_path <- here::here(
  "data", "clinical", "protocol_2101175",
  "ProtocolNo2101175PRM_DATA_2026-05-18_2135.csv"
)
raw <- read_csv(data_path, show_col_types = FALSE)
ctcae_term_map <- load_ctcae_term_labels()

# ---- Factor mappings (from REDCap companion script) ----
# Safe recode: never assign a full vector into is.na(out) slots (avoids label corruption).
map_chr <- function(x, mapping) {
  x_chr <- as.character(x)
  out <- unname(mapping[x_chr])
  unmatched <- is.na(out) & !is.na(x) & x_chr != ""
  if (any(unmatched)) {
    out[unmatched] <- paste0("Code ", x_chr[unmatched])
  }
  out[is.na(x) | x_chr == ""] <- NA_character_
  out
}

# Stop knit if categorical table cells do not sum to the stated total N.
assert_partition_sums <- function(table_name, component_counts, total_n) {
  comp <- as.integer(component_counts)
  if (any(is.na(comp))) {
    stop(table_name, ": missing component counts (NA).", call. = FALSE)
  }
  s <- sum(comp)
  if (s != total_n) {
    stop(
      sprintf(
        "%s: components (%s) sum to %d but total N = %d",
        table_name,
        paste(comp, collapse = " + "),
        s,
        total_n
      ),
      call. = FALSE
    )
  }
  invisible(TRUE)
}

cohort_lab <- c(
  "1" = "Cohort 1 low dose (0.6 mg BID)",
  "2" = "Cohort 1 high dose (0.6 mg TID)",
  "3" = "Cohort 2 (0.6 mg BID + nivolumab)"
)
gender_lab <- c("1" = "Male", "2" = "Female", "3" = "Other")
race_lab <- c(
  "1" = "Black", "2" = "White", "3" = "Asian", "4" = "American Indian",
  "5" = "Hawaiian/Pacific Islander", "6" = "Other/Unknown"
)
ethnicity_lab <- c(
  "1" = "Hispanic or Latino", "2" = "NOT Hispanic or Latino", "3" = "Unknown / Not Reported"
)
ecog_lab <- c(
  "0" = "0", "1" = "1", "2" = "2", "3" = "3", "4" = "4", "-99" = "Not collected"
)
enrolled_lab <- c("0" = "No", "1" = "Yes")
status_lab <- c("0" = "Alive", "1" = "Dead", "9" = "Lost to follow-up")
ae_grade_lab <- c(
  "1" = "1", "2" = "2", "3" = "3", "4" = "4", "5" = "5"
)
sae_lab <- c("0" = "No", "1" = "Yes")
attr_lab <- c(
  "1" = "Unrelated", "2" = "Unlikely", "3" = "Possible", "4" = "Probable", "5" = "Definite", "6" = "N/A"
)

dat <- raw %>%
  mutate(
    study_id = as.integer(study_id),
    cohort_label = map_chr(cohort, cohort_lab),
    gender_label = map_chr(gender, gender_lab),
    race_label = map_chr(race, race_lab),
    ethnicity_label = map_chr(ethnicity, ethnicity_lab),
    enrolled_label = map_chr(was_the_subject_enrolled, enrolled_lab),
    ecog_label = map_chr(ecog_performance_status, ecog_lab),
    subject_status_label = map_chr(subject_status_2, status_lab),
    ae_grade_label = map_chr(aegrade, ae_grade_lab),
    sae_label = map_chr(is_this_a_sae, sae_lab),
    ae_attribution_label = map_chr(ae_attribution_colchicine, attr_lab),
    crp_value = suppressWarnings(as.numeric(c_reactive_protein_value)),
    ferritin_value_num = suppressWarnings(as.numeric(ferritin_value)),
    esr_value_num = suppressWarnings(as.numeric(erythrocyte_sedimentation_value)),
    date_of_birth_dt = suppressWarnings(ymd(date_of_birth)),
    date_enrollment_dt = suppressWarnings(ymd(date_of_enrollment)),
    date_final_status_dt = suppressWarnings(ymd(date_of_final_status)),
    treatment_start_dt = suppressWarnings(ymd(treatment_start_date))
  )

# ---- Cohort / visit helpers ----
baseline_events <- c(
  "cycle_1_day_1__3_d_arm_1", "cycle_1_day_1_arm_2", "cycle_1_day_1_arm_3"
)
post_baseline_events <- c(
  "cycle_1_day_15__3_arm_1",
  "cycle_1_day_8_arm_2", "cycle_1_day_15_arm_2",
  "cycle_1_day_8_arm_3", "cycle_1_day_15_arm_3"
)
screening_events <- c("screening_28_days_arm_1", "screening_arm_2", "screening_arm_3")

screen_disposition <- build_protocol2101175_screen_disposition(dat)
consented_ids <- screen_disposition %>%
  filter(!is.na(consent_date)) %>%
  pull(study_id) %>%
  sort()
enrolled_ids <- screen_disposition %>%
  filter(enrolled) %>%
  pull(study_id) %>%
  sort()
screen_fail_ids <- screen_disposition %>%
  filter(screen_fail) %>%
  pull(study_id) %>%
  sort()

enrolled_original_ids <- screen_disposition %>%
  filter(enrolled, protocol_era == "original") %>%
  pull(study_id) %>%
  sort()
enrolled_amended_ids <- screen_disposition %>%
  filter(enrolled, protocol_era == "amended") %>%
  pull(study_id) %>%
  sort()
km_eligible_ids <- protocol2101175_km_eligible_ids(screen_disposition)

consented_n <- length(consented_ids)
enrolled_n <- length(enrolled_ids)
enrolled_original_n <- length(enrolled_original_ids)
enrolled_amended_n <- length(enrolled_amended_ids)
km_eligible_n <- length(km_eligible_ids)
screen_fail_n <- length(screen_fail_ids)
screen_fail_original_n <- screen_disposition %>%
  filter(screen_fail, protocol_era == "original") %>%
  nrow()
screen_fail_amended_n <- screen_disposition %>%
  filter(screen_fail, protocol_era == "amended") %>%
  nrow()
screened_n <- consented_n

assert_partition_sums(
  "Accrual disposition (consented vs enrolled + screen-fail)",
  c(enrolled_n, screen_fail_n),
  consented_n
)
assert_partition_sums(
  "Enrolled patients by protocol schedule",
  c(enrolled_original_n, enrolled_amended_n),
  enrolled_n
)
assert_partition_sums(
  "Screen failures by protocol schedule",
  c(screen_fail_original_n, screen_fail_amended_n),
  screen_fail_n
)

screen_fail_summary <- summarise_screen_fail_reasons(screen_disposition)

# ---- Primary endpoint: max % CRP decline ----
crp_long <- dat %>%
  filter(study_id %in% enrolled_ids, !is.na(crp_value)) %>%
  select(study_id, redcap_event_name, crp_value, cohort_label) %>%
  distinct()

derive_crp_endpoint <- function(id, d_long) {
  sub <- d_long %>% filter(study_id == id)
  base_row <- sub %>%
    filter(redcap_event_name %in% baseline_events) %>%
    arrange(redcap_event_name) %>%
    slice(1)
  if (nrow(base_row) == 0) {
    base_row <- sub %>%
      filter(redcap_event_name %in% screening_events) %>%
      arrange(redcap_event_name) %>%
      slice(1)
  }
  if (nrow(base_row) == 0) return(NULL)
  baseline <- base_row$crp_value[1]
  if (is.na(baseline) || baseline <= 0) return(NULL)
  post <- sub %>%
    filter(redcap_event_name %in% post_baseline_events) %>%
    pull(crp_value)
  if (length(post) == 0) return(NULL)
  # Decline cannot be negative by definition; increases contribute 0% decline.
  pct_declines <- pmax(100 * (baseline - post) / baseline, 0)
  tibble(
    study_id = id,
    baseline_crp_mg_l = baseline,
    min_on_tx_crp_mg_l = min(post),
    max_pct_decline = max(pct_declines),
    n_crp_on_tx = length(post)
  )
}

crp_endpoint <- bind_rows(lapply(enrolled_ids, derive_crp_endpoint, crp_long))
evaluable_efficacy_n <- nrow(crp_endpoint)
evaluable_toxicity_n <- enrolled_n

# Primary one-sided t-test (protocol: alpha = 0.025 for Cohort 1 low dose)
primary_test <- if (evaluable_efficacy_n >= 2) {
  t.test(crp_endpoint$max_pct_decline, mu = 0, alternative = "greater")
} else {
  NULL
}

primary_mean <- mean(crp_endpoint$max_pct_decline, na.rm = TRUE)
primary_sd <- sd(crp_endpoint$max_pct_decline, na.rm = TRUE)
primary_p <- if (!is.null(primary_test)) primary_test$p.value else NA_real_
primary_t <- if (!is.null(primary_test)) primary_test$statistic else NA_real_
primary_ci_low <- if (!is.null(primary_test)) primary_test$conf.int[1] else NA_real_

cohens_d_primary <- if (evaluable_efficacy_n >= 2 && !is.na(primary_sd) && primary_sd > 0) {
  primary_mean / primary_sd
} else {
  NA_real_
}

posthoc_power_pct <- if (evaluable_efficacy_n >= 2 && !is.na(cohens_d_primary)) {
  n <- evaluable_efficacy_n
  ncp <- cohens_d_primary * sqrt(n)
  t_crit <- qt(1 - 0.025, df = n - 1)
  (1 - pt(t_crit, df = n - 1, ncp = ncp)) * 100
} else {
  NA_real_
}

# ---- Patient-level master (one row per enrolled subject) ----
# Demographics from screening visit; first non-missing value per field (not first table row).
patient_master <- tibble(study_id = enrolled_ids) %>%
  left_join(
    dat %>%
      filter(study_id %in% enrolled_ids, redcap_event_name %in% screening_events) %>%
      group_by(study_id) %>%
      summarise(
        gender_code = suppressWarnings(as.integer(first(na.omit(gender)))),
        race_code = suppressWarnings(as.integer(first(na.omit(race)))),
        ethnicity_code = suppressWarnings(as.integer(first(na.omit(ethnicity)))),
        ecog_code = suppressWarnings(as.integer(first(na.omit(ecog_performance_status)))),
        date_of_birth_dt = first(na.omit(date_of_birth_dt)),
        date_enrollment_dt = first(na.omit(date_enrollment_dt)),
        cohort_code = suppressWarnings(as.integer(first(na.omit(cohort)))),
        .groups = "drop"
      ),
    by = "study_id"
  ) %>%
  mutate(
    gender_label = map_chr(gender_code, gender_lab),
    race_label = map_chr(race_code, race_lab),
    ethnicity_label = map_chr(ethnicity_code, ethnicity_lab),
    ecog_label = map_chr(ecog_code, ecog_lab),
    cohort_label = map_chr(cohort_code, cohort_lab),
    age_years = if_else(
      !is.na(date_of_birth_dt) & !is.na(date_enrollment_dt),
      as.numeric(interval(date_of_birth_dt, date_enrollment_dt) / years(1)),
      NA_real_
    )
  )

if (nrow(patient_master) != enrolled_n) {
  stop("patient_master row count does not match enrolled_n.", call. = FALSE)
}

baseline <- patient_master

# Prior therapies → therapy class (repeat instrument)
prior_tx <- dat %>%
  filter(
    study_id %in% enrolled_ids,
    redcap_repeat_instrument == "prior_treatment_therapy",
    prior_treatment_therapies != ""
  ) %>%
  distinct(study_id, prior_treatment_therapies, therapy_start_date) %>%
  mutate(therapy_class = classify_prior_therapy(prior_treatment_therapies))

therapy_class_levels <- c(
  "Chemotherapy",
  "Immunotherapy",
  "Targeted therapy",
  "Radiation therapy",
  "Surgery",
  "Other"
)

therapy_class_levels_table1 <- c(
  "Chemotherapy",
  "Immunotherapy",
  "Targeted therapy",
  "Radiation therapy",
  "Surgery"
)

prior_tx_by_patient <- prior_tx %>%
  distinct(study_id, therapy_class) %>%
  filter(!is.na(therapy_class)) %>%
  mutate(received = TRUE) %>%
  pivot_wider(
    names_from = therapy_class,
    values_from = received,
    values_fill = list(received = FALSE)
  )

for (col in therapy_class_levels) {
  if (!col %in% names(prior_tx_by_patient)) {
    prior_tx_by_patient[[col]] <- FALSE
  }
}

prior_tx_summary <- prior_tx %>%
  group_by(therapy_class) %>%
  summarise(
    `N patients` = n_distinct(study_id),
    `N regimen entries` = n(),
    .groups = "drop"
  ) %>%
  arrange(desc(`N patients`))

# Primary tumor type (inferred from prior therapy / surgery text in REDCap)
tumor_by_patient <- prior_tx %>%
  group_by(study_id) %>%
  summarise(
    primary_tumor_type = infer_primary_tumor_type(
      study_id[1],
      prior_treatment_therapies
    ),
    .groups = "drop"
  )

patient_master <- patient_master %>%
  left_join(tumor_by_patient, by = "study_id")

# Systemic therapy line count approximation:
# collapse same-date combinations into one regimen line and exclude RT/surgery/other.
prior_systemic_lines <- prior_tx %>%
  filter(therapy_class %in% c("Chemotherapy", "Immunotherapy", "Targeted therapy")) %>%
  mutate(systemic_start_date = suppressWarnings(ymd(therapy_start_date))) %>%
  group_by(study_id) %>%
  summarise(
    n_systemic_lines = if_else(
      any(!is.na(systemic_start_date)),
      n_distinct(systemic_start_date[!is.na(systemic_start_date)]),
      n_distinct(prior_treatment_therapies)
    ),
    .groups = "drop"
  ) %>%
  right_join(tibble(study_id = enrolled_ids), by = "study_id") %>%
  mutate(n_systemic_lines = replace_na(n_systemic_lines, 0L))

tumor_type_summary <- patient_master %>%
  count(primary_tumor_type, name = "N") %>%
  arrange(desc(N)) %>%
  mutate(
    `Cancer type (patient-level)` = paste0(primary_tumor_type, " (n = ", N, ")")
  )

assert_partition_sums(
  "Patient characteristics: primary cancer type",
  tumor_type_summary$N,
  enrolled_n
)

# ---- Adverse events ----
# TRAE-only: attribution to colchicine is Possible / Probable / Definite.
ae_all <- dat %>%
  filter(
    study_id %in% enrolled_ids,
    ae_term_ctcaev5 != "",
    ae_attribution_colchicine %in% c(3, 4, 5)
  ) %>%
  mutate(
    ae_grade_num = suppressWarnings(as.numeric(aegrade)),
    ae_term_display = ctcae_term_label(ae_term_ctcaev5, ctcae_term_map),
    ae_other_text = if ("if_you_entered_other_pleas" %in% names(.))
      trimws(as.character(if_you_entered_other_pleas))
    else
      NA_character_
  ) %>%
  mutate(
    ae_term_display = if_else(
      !is.na(ae_other_text) & ae_other_text != "" &
        (is.na(ae_term_display) |
          stringr::str_detect(tolower(ae_term_display), "^other|not elsewhere classified")),
      ae_other_text,
      ae_term_display
    )
  )

ae_by_term_grade <- ae_all %>%
  count(ae_term_display, ae_grade_label, name = "n") %>%
  arrange(desc(n))

# Worst CTCAE grade per patient per preferred term (do not count lower grades separately).
ae_worst_per_patient <- ae_all %>%
  filter(!is.na(ae_grade_num)) %>%
  group_by(study_id, ae_term_display) %>%
  slice_max(ae_grade_num, n = 1, with_ties = FALSE) %>%
  ungroup()

ae_by_term_grade_patient <- ae_worst_per_patient %>%
  count(ae_term_display, ae_grade_label, name = "N patients") %>%
  mutate(`N (%)` = paste0(`N patients`, " (", sprintf("%.1f", 100 * `N patients` / enrolled_n), "%)")) %>%
  arrange(desc(`N patients`), ae_term_display, suppressWarnings(as.numeric(ae_grade_label)))

ae_any_grade <- ae_all %>%
  distinct(study_id, ae_term_display) %>%
  count(study_id, name = "n_aes") %>%
  right_join(tibble(study_id = enrolled_ids), by = "study_id") %>%
  mutate(n_aes = replace_na(n_aes, 0L))

grade3plus_rate <- ae_worst_per_patient %>%
  filter(ae_grade_num >= 3) %>%
  distinct(study_id) %>%
  nrow() / enrolled_n * 100

trae_patient_n <- ae_all %>%
  distinct(study_id) %>%
  nrow()
trae_patient_pct <- 100 * trae_patient_n / enrolled_n

trae_diarrhea_n <- ae_worst_per_patient %>%
  filter(tolower(ae_term_display) == "diarrhea") %>%
  distinct(study_id) %>%
  nrow()
trae_diarrhea_pct <- 100 * trae_diarrhea_n / enrolled_n

# Helpers for inline text (also used in survival tables)
fmt_p <- function(p) {
  if (is.na(p)) return("NA")
  if (p < 0.001) return("<0.001")
  format(round(p, 3), nsmall = 3)
}
fmt_num <- function(x, digits = 1) format(round(x, digits), nsmall = digits)

.tbl_n <- 0L
.fig_n <- 0L
tbl_cap <- function(title) {
  .tbl_n <<- .tbl_n + 1L
  paste0("Table ", .tbl_n, ". ", title)
}
fig_cap <- function(title) {
  .fig_n <<- .fig_n + 1L
  paste0("Figure ", .fig_n, ". ", title)
}

# ---- Survival / status (PFS/OS derived from RECIST, discontinuation, final status) ----
# Restrict to original-schedule enrollments with serial imaging surveillance (Arm 1).
surv_dat <- build_protocol2101175_survival(dat, km_eligible_ids)

# Best % change in target-lesion SOD (exploratory; patients with ≥2 scan sets)
tumor_change <- build_protocol2101175_tumor_change(dat, enrolled_ids)

status_tbl <- surv_dat %>%
  transmute(
    study_id,
    status = coalesce(final_status, "Not recorded"),
    date_final = if_else(os_event, death_date, os_end_date)
  )

n_dead <- sum(surv_dat$os_event, na.rm = TRUE)
n_pfs_events <- sum(surv_dat$pfs_event, na.rm = TRUE)

surv_pfs <- surv_dat %>%
  filter(!is.na(pfs_months), pfs_months >= 0)

surv_os <- surv_dat %>%
  filter(!is.na(os_months), os_months >= 0)

fit_pfs <- if (nrow(surv_pfs) >= 3 && sum(surv_pfs$pfs_event) >= 1) {
  survfit(Surv(pfs_months, pfs_event) ~ 1, data = surv_pfs)
} else {
  NULL
}

fit_os <- if (nrow(surv_os) >= 3 && sum(surv_os$os_event) >= 1) {
  survfit(Surv(os_months, os_event) ~ 1, data = surv_os)
} else {
  NULL
}

median_surv_months <- function(fit) {
  if (is.null(fit)) return(list(median = NA_real_, lcl = NA_real_, ucl = NA_real_))
  s <- summary(fit)$table
  if (is.matrix(s)) {
    list(
      median = unname(s[1, "median"]),
      lcl = unname(s[1, "0.95LCL"]),
      ucl = unname(s[1, "0.95UCL"])
    )
  } else {
    list(median = NA_real_, lcl = NA_real_, ucl = NA_real_)
  }
}

pfs_med <- median_surv_months(fit_pfs)
os_med <- median_surv_months(fit_os)

pfs_os_tbl <- surv_dat %>%
  transmute(
    `Study ID` = study_id,
    `Index date` = as.character(time_zero),
    `Best RECIST on study` = coalesce(best_recist, "Not assessed"),
    `PFS event` = if_else(pfs_event, "Progression", "Censored"),
    `PFS date` = as.character(pfs_end_date),
    `PFS (months)` = fmt_num(pfs_months, 1),
    `OS event` = if_else(os_event, "Death", "Censored"),
    `OS date` = as.character(os_end_date),
    `OS (months)` = fmt_num(os_months, 1),
    `Final status` = coalesce(final_status, "Not recorded")
  )

ctr_table_style <- function(kbl) {
  kbl %>%
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) %>%
    row_spec(0, bold = TRUE, background = "#D9E2F3")
}
```

<div id="inline-edit-toolbar" style="position:fixed;top:10px;right:10px;z-index:9999;background:#ffffff;border:1px solid #cfd8dc;border-radius:8px;padding:8px;box-shadow:0 2px 10px rgba(0,0,0,0.12);font-family:Arial,sans-serif;">
  <button id="toggle-inline-edit" style="margin-right:8px;">Enable Edit Mode</button>
  <button id="download-inline-html" disabled style="margin-right:8px;">Download Edited HTML</button>
  <button id="clear-inline-edits" title="Clear browser-saved edits for this report">Clear Saved Edits</button>
</div>

<script>
(function () {
  var editOn = false;
  var root = document.querySelector(".main-container") || document.body;
  var toggleBtn = document.getElementById("toggle-inline-edit");
  var downloadBtn = document.getElementById("download-inline-html");
  var clearBtn = document.getElementById("clear-inline-edits");
  var toolbar = document.getElementById("inline-edit-toolbar");
  var storageKey = "ctr_inline_edits:" + location.pathname;
  var saveTimer = null;

  function saveEdits() {
    try {
      localStorage.setItem(storageKey, root.innerHTML);
    } catch (e) {
      console.warn("Could not save local edits:", e);
    }
  }

  function scheduleSave() {
    if (!editOn) return;
    clearTimeout(saveTimer);
    saveTimer = setTimeout(saveEdits, 250);
  }

  function setEditState(next) {
    editOn = next;
    if (editOn) {
      root.setAttribute("contenteditable", "true");
      root.setAttribute("spellcheck", "true");
      root.style.outline = "2px dashed #4a90e2";
      root.style.outlineOffset = "6px";
      toggleBtn.textContent = "Disable Edit Mode";
      downloadBtn.disabled = false;
    } else {
      root.removeAttribute("contenteditable");
      root.style.outline = "";
      root.style.outlineOffset = "";
      toggleBtn.textContent = "Enable Edit Mode";
      downloadBtn.disabled = true;
    }
  }

  function downloadEditedHtml() {
    var prior = toolbar ? toolbar.style.display : "";
    if (toolbar) toolbar.style.display = "none";
    var html = "<!doctype html>\n" + document.documentElement.outerHTML;
    if (toolbar) toolbar.style.display = prior;

    var blob = new Blob([html], { type: "text/html;charset=utf-8" });
    var url = URL.createObjectURL(blob);
    var a = document.createElement("a");
    a.href = url;
    a.download = "CTR_Submission_Results_Protocol2101175_edited.html";
    document.body.appendChild(a);
    a.click();
    a.remove();
    URL.revokeObjectURL(url);
  }

  toggleBtn.addEventListener("click", function () {
    setEditState(!editOn);
  });

  downloadBtn.addEventListener("click", downloadEditedHtml);
  clearBtn.addEventListener("click", function () {
    localStorage.removeItem(storageKey);
    alert("Saved local edits cleared for this report URL.");
  });

  document.addEventListener("keydown", function (e) {
    if (editOn && (e.metaKey || e.ctrlKey) && e.key.toLowerCase() === "s") {
      e.preventDefault();
      downloadEditedHtml();
    }
  });

  root.addEventListener("input", scheduleSave);

  try {
    var saved = localStorage.getItem(storageKey);
    if (saved && saved.length > 0) {
      root.innerHTML = saved;
    }
  } catch (e) {
    console.warn("Could not restore local edits:", e);
  }
})();
</script>

# Clinical Trial Results

**Protocol:** STUDY-21-001175 (IND 156493), version 3.0 (22 May 2023)  
**Institution:** Icahn School of Medicine at Mount Sinai / Mount Sinai Hospital  
**Principal investigator:** Deborah Y. Doroshow, MD, PhD  
**Data cut:** `r format(Sys.Date(), "%d %B %Y")` (REDCap export `ProtocolNo2101175PRM_DATA_2026-05-18_2135.csv`)

---

## Abstract

**Background:** Pro-tumorigenic inflammation in advanced solid tumors is linked to IL-1β–NLRP3 signaling and elevated C-reactive protein (CRP). We conducted a single-center open-label pilot trial to evaluate colchicine effects on peripheral blood CRP in patients with advanced or recurrent solid tumors.

**Methods:** Eligible patients had metastatic or recurrent solid tumors with CRP greater than 5 mg/L and received colchicine 0.6 mg orally twice daily (Cohort 1 low dose). A protocol amendment limited later enrollments to a fixed 2-week course; earlier enrollments received indefinite colchicine with serial imaging. The primary endpoint was maximum percent CRP decline from cycle 1 day 1 to the lowest on-treatment value. Planned high-dose and adjuvant cohorts did not accrue.

**Results:** `r consented_n` patients consented; `r enrolled_n` enrolled and `r screen_fail_n` screen-failed. `r enrolled_original_n` were treated under the original schedule and `r enrolled_amended_n` under the amended schedule. `r evaluable_efficacy_n` patients were evaluable for the primary endpoint. Mean maximum CRP decline was `r fmt_num(primary_mean)`% (SD `r fmt_num(primary_sd)`%); one-sided one-sample *t*-test *p* = `r fmt_p(primary_p)`, not meeting prespecified α = 0.025. Treatment-related adverse events occurred in `r trae_patient_n` of `r enrolled_n` patients (`r fmt_num(trae_patient_pct)`%); all were grade 1–2, most commonly diarrhea (`r trae_diarrhea_n`/`r enrolled_n`, `r fmt_num(trae_diarrhea_pct)`%).

**Conclusion:** Colchicine 0.6 mg twice daily was tolerable with modest CRP suppression but did not meet the prespecified primary endpoint. The data do not support colchicine as a re-sensitization strategy in this setting. Trial registration: ClinicalTrials.gov, NCT05279690.

**Key words:** colchicine; C-reactive protein; inflammation; solid tumors; pilot trial

---

## Lessons Learned

- **Key outcome:** The prespecified primary CRP endpoint was not met (one-sided *p* = `r fmt_p(primary_p)` vs α = 0.025). With only `r evaluable_efficacy_n` of 15 planned evaluable patients, the trial was underpowered to confirm the observed `r fmt_num(primary_mean)`% mean CRP decline at the prespecified significance threshold.
- **Key lesson:** Short-course low-dose colchicine was tolerable but yielded only a modest pharmacodynamic signal; future inflammation-modulation pilots in advanced solid tumors will likely need narrower disease selection, less restrictive eligibility, or larger cohorts to draw definitive conclusions.
- **Key reason:** Accrual lagged because eligible patients were difficult to identify in routine oncology clinics (CRP greater than 5 mg/L plus organ-function requirements), `r screen_fail_n` of `r consented_n` consented patients (`r fmt_num(100 * screen_fail_n / consented_n, 1)`%) screen-failed before treatment (including 2 patient withdrawals, 1 organ dysfunction, and 1 acute clinical deterioration), and the adjuvant urothelial carcinoma cohort closed with zero enrollments.

---

## Trial Information

```{r trial-info-table}
trial_info <- tibble(
  Field = c(
    "Disease",
    "Stage of disease / treatment",
    "Prior therapy",
    "Type of study",
    "Primary endpoint",
    "Secondary endpoints",
    "Additional details"
  ),
  Description = c(
    "Advanced/recurrent solid tumors (Cohort 1); resected high-risk urothelial carcinoma planned for adjuvant nivolumab (Cohort 2, closed to accrual)",
    "Metastatic/recurrent solid tumors (Cohort 1); post-surgical high-risk urothelial carcinoma (Cohort 2, not enrolled)",
    "Variable prior systemic and local therapy per disease (see Patient Characteristics)",
    "Single-center, open-label, non-randomized pilot",
    "Maximum percentage decline in peripheral blood CRP from cycle 1 day 1 during on-study colchicine",
    "CTCAE v5.0 safety; PFS (Cohort 1); DFS/OS (Cohort 2); exploratory cytokines, ctDNA, tissue correlatives",
    paste0(
      "Cohort 1 low-dose (BID) and high-dose (TID) were sequential; only low-dose enrolled. ",
      "Cohort 2 closed for poor accrual before enrollment. During conduct, the protocol was amended: ",
      "the first ", enrolled_original_n, " enrolled patients received indefinite on-study colchicine ",
      "with serial imaging surveillance; the subsequent ", enrolled_amended_n,
      " enrolled patients received a fixed 2-week colchicine course (phase 0 pharmacodynamic design). ",
      "Exploratory Kaplan–Meier PFS/OS analyses were restricted to the original-schedule cohort (n = ",
      km_eligible_n, ")."
    )
  )
)
trial_info %>%
  kbl(
    caption = tbl_cap("Trial information."),
    col.names = c("Trial information", "Description"),
    align = c("l", "l")
  ) %>%
  ctr_table_style()
```

---

## Drug Information

```{r drug-info-table}
drug_info <- tibble(
  `Drug information` = c(
    "Generic/working name", "Drug type", "Drug class", "Dose (Cohort 1 low)",
    "Dose (Cohort 1 high)", "Dose (Cohort 2)", "Route", "Schedule"
  ),
  Details = c(
    "Colchicine", "Anti-inflammatory alkaloid", "Microtubule polymerization inhibitor (indirect NLRP3 pathway modulation)",
    "0.6 mg oral BID × 14 days", "0.6 mg oral TID × 14 days (not enrolled)",
    "0.6 mg oral BID × 28 days with nivolumab 480 mg IV every 4 weeks (cohort closed)",
    "Oral", "One cycle per protocol; missed doses not doubled"
  )
)
drug_info %>%
  kbl(
    caption = tbl_cap("Drug information."),
    col.names = c("Drug information", "Details"),
    align = c("l", "l")
  ) %>%
  ctr_table_style()
```

---

## Primary Assessment Method

```{r primary-assessment-table}
primary_assess <- tibble(
  `Primary assessment` = c(
    "Title",
    "Number of patients who signed informed consent",
    "Number of patients enrolled",
    "Number of screen failures before enrollment",
    "Number evaluable for toxicity",
    "Number evaluable for primary efficacy (CRP)",
    "Evaluation method",
    "Primary analysis result (Cohort 1 low dose)"
  ),
  Value = c(
    "Maximum percent decline in CRP from baseline (cycle 1 day 1) during post-baseline C1 on-treatment visits",
    as.character(consented_n),
    as.character(enrolled_n),
    as.character(screen_fail_n),
    as.character(evaluable_toxicity_n),
    as.character(evaluable_efficacy_n),
    "Peripheral blood CRP (mg/L) at protocol C1 visits; baseline = cycle 1 day 1 (or screening fallback if missing); minimum post-baseline value at C1D8/C1D15 used for maximum decline. Cycle 2 values excluded from primary endpoint per protocol treatment window.",
    paste0(
      "Mean max decline ", fmt_num(primary_mean), "% (SD ", fmt_num(primary_sd), "%); ",
      "one-sided one-sample t-test vs 0, t = ", fmt_num(primary_t, 2), ", p = ", fmt_p(primary_p),
      " (prespecified α = 0.025)"
    )
  )
)
primary_assess %>%
  kbl(
    caption = tbl_cap("Primary assessment method."),
    col.names = c("Primary assessment", "Value"),
    align = c("l", "l")
  ) %>%
  ctr_table_style()
```

---

## Accrual Disposition

All `r consented_n` subjects who signed informed consent underwent screening evaluation. Enrollment status was recorded on the REDCap screening form (`was_the_subject_enrolled`: 1 = enrolled, 0 = screen failure). **During trial conduct, a protocol amendment limited on-study colchicine to a fixed 2-week course for patients screened under REDCap Arm 2; patients screened under Arm 1 received the original indefinite on-study schedule with serial imaging surveillance.** `r enrolled_original_n` enrolled patients are Kaplan–Meier–evaluable under the original schedule; `r enrolled_amended_n` enrolled after the amendment contributed to toxicity and primary CRP pharmacodynamic analyses but were excluded from Kaplan–Meier PFS/OS estimates. When organ-function or CRP thresholds were not met, the report states the clinical failure (e.g., inadequate organ function) and the screening laboratory value when available. REDCap may classify these as “inclusion criteria” screen failures because adequate organ function and CRP elevation are protocol inclusion requirements.

```{r accrual-disposition-table}
accrual_overview <- tibble(
  Disposition = c(
    "Signed informed consent",
    "Enrolled on study",
    "  Original schedule (indefinite colchicine + surveillance imaging)",
    "  Amended schedule (fixed 2-week phase 0 intervention)",
    "Screen failure before enrollment",
    "  Before original schedule",
    "  After protocol amendment"
  ),
  N = c(
    consented_n,
    enrolled_n,
    enrolled_original_n,
    enrolled_amended_n,
    screen_fail_n,
    screen_fail_original_n,
    screen_fail_amended_n
  )
) %>%
  mutate(`N (%)` = case_when(
    Disposition == "Signed informed consent" ~ as.character(N),
    Disposition == "Enrolled on study" ~ paste0(N, " (", trimws(fmt_num(100 * N / consented_n, 1)), "% of consented)"),
    Disposition == "Screen failure before enrollment" ~ paste0(N, " (", trimws(fmt_num(100 * N / consented_n, 1)), "% of consented)"),
    grepl("^  Original schedule", Disposition) ~ paste0(N, " (Kaplan–Meier–evaluable)"),
    grepl("^  Amended schedule", Disposition) ~ paste0(N, " (primary CRP / toxicity only)"),
    TRUE ~ paste0(N, " (", trimws(fmt_num(100 * N / consented_n, 1)), "% of consented)")
  ))

accrual_overview %>%
  kbl(
    caption = tbl_cap("Accrual disposition among consented subjects."),
    col.names = c("Disposition", "N", "N (%)"),
    align = c("l", "r", "r")
  ) %>%
  ctr_table_style()

if (nrow(screen_fail_summary) > 0) {
  screen_fail_summary %>%
    mutate(
      `N (%)` = paste0(`N`, " (", trimws(fmt_num(100 * `N` / consented_n, 1)), "% of consented)")
    ) %>%
    select(`Screen-failure reason`, N, `N (%)`, Details) %>%
    kbl(
      caption = tbl_cap("Reasons for screen failure before enrollment."),
      col.names = c("Screen-failure reason", "N", "N (%)", "Details"),
      align = c("l", "r", "r", "l")
    ) %>%
    ctr_table_style()
}

if (screen_fail_n > 0) {
  screen_disposition %>%
    filter(screen_fail) %>%
    transmute(
      `Consent date` = format(consent_date, "%Y-%m-%d"),
      `Screen-failure reason` = screen_fail_reason,
      Details = screen_fail_detail
    ) %>%
    kbl(
      caption = tbl_cap("Screen-failure subjects: reason and supporting detail."),
      col.names = c("Consent date", "Screen-failure reason", "Details"),
      align = c("l", "l", "l")
    ) %>%
    ctr_table_style()
}
```

---

## Patient Characteristics

```{r patient-chars-table}
n_male <- sum(patient_master$gender_label == "Male", na.rm = TRUE)
n_female <- sum(patient_master$gender_label == "Female", na.rm = TRUE)
n_sex_other <- sum(patient_master$gender_label == "Other", na.rm = TRUE)
n_sex_unknown <- sum(is.na(patient_master$gender_label), na.rm = TRUE)

assert_partition_sums(
  "Patient characteristics: sex",
  c(n_male, n_female, n_sex_other, n_sex_unknown),
  enrolled_n
)

ecog01 <- sum(patient_master$ecog_code %in% c(0, 1), na.rm = TRUE)
ecog2p <- sum(patient_master$ecog_code %in% c(2, 3, 4), na.rm = TRUE)
ecog_unknown <- enrolled_n - ecog01 - ecog2p

assert_partition_sums(
  "Patient characteristics: ECOG",
  c(ecog01, ecog2p, ecog_unknown),
  enrolled_n
)

age_vec <- patient_master$age_years[!is.na(patient_master$age_years)]
age_med <- if (length(age_vec)) median(age_vec) else NA_real_
age_rng <- if (length(age_vec)) range(age_vec) else c(NA_real_, NA_real_)
n_age_known <- length(age_vec)

prior_med <- median(prior_systemic_lines$n_systemic_lines, na.rm = TRUE)
prior_rng <- range(prior_systemic_lines$n_systemic_lines, na.rm = TRUE)

fmt_n_pct <- function(n, d = enrolled_n) paste0(n, " (", trimws(fmt_num(100 * n / d, 1)), "%)")

tumor_type_rows <- tumor_type_summary %>%
  mutate(
    `Patient characteristics` = paste0("  ", primary_tumor_type, ", n (%)"),
    Value = fmt_n_pct(N)
  ) %>%
  select(`Patient characteristics`, Value)

therapy_class_rows <- tibble(therapy_class = therapy_class_levels_table1) %>%
  left_join(
    prior_tx %>%
      group_by(therapy_class) %>%
      summarise(`N patients` = n_distinct(study_id), .groups = "drop"),
    by = "therapy_class"
  ) %>%
  mutate(`N patients` = replace_na(`N patients`, 0L)) %>%
  mutate(
    `Patient characteristics` = paste0("  ", therapy_class, ", n (%)"),
    Value = fmt_n_pct(`N patients`)
  ) %>%
  select(`Patient characteristics`, Value)

pat_chars <- bind_rows(
  tibble(
    `Patient characteristics` = c(
      "Number enrolled (Cohort 1 low dose)",
      "Sex, male, n (%)",
      "Sex, female, n (%)",
      "Age at enrollment, median (range), years",
      "Performance status ECOG 0–1, n (%)",
      "Performance status ECOG ≥2, n (%)",
      "Prior systemic therapy lines, median (range), n"
    ),
    Value = c(
      as.character(enrolled_n),
      fmt_n_pct(n_male),
      fmt_n_pct(n_female),
      if (n_age_known == 0) {
        "Not available"
      } else {
        paste0(
          fmt_num(age_med, 0), " (", fmt_num(age_rng[1], 0), "–", fmt_num(age_rng[2], 0), ")"
        )
      },
      fmt_n_pct(ecog01),
      fmt_n_pct(ecog2p),
      if (length(prior_med) == 0 || is.na(prior_med)) {
        "Not available"
      } else {
        paste0(prior_med, " (", prior_rng[1], "–", prior_rng[2], ")")
      }
    )
  ),
  tibble(`Patient characteristics` = "Exact tumor type", Value = ""),
  tumor_type_rows,
  tibble(`Patient characteristics` = "Prior therapy class (N patients)", Value = ""),
  therapy_class_rows
)
pat_chars %>%
  kbl(
    caption = tbl_cap("Patient characteristics (Cohort 1 low dose)."),
    col.names = c("Characteristic", "Value"),
    align = c("l", "l")
  ) %>%
  ctr_table_style()
```

---

## Primary Assessment Results

### CRP endpoint by patient

```{r crp-endpoint-table}
crp_endpoint %>%
  mutate(
    baseline_crp_mg_l = round(baseline_crp_mg_l, 1),
    min_on_tx_crp_mg_l = round(min_on_tx_crp_mg_l, 1),
    max_pct_decline = round(max_pct_decline, 1)
  ) %>%
  rename(
    `Study ID` = study_id,
    `Baseline CRP (mg/L)` = baseline_crp_mg_l,
    `Lowest on-treatment CRP (mg/L)` = min_on_tx_crp_mg_l,
    `Max % decline` = max_pct_decline,
    `N on-tx CRP values` = n_crp_on_tx
  ) %>%
  kbl(caption = tbl_cap("Primary CRP endpoint derivation by enrolled patient.")) %>%
  ctr_table_style()
```

```{r waterfall-plot, fig.cap=fig_cap("Maximum percent decline in C-reactive protein (CRP) from baseline during on-treatment assessment by enrolled patient (Cohort 1 low dose). Abbreviations: CRP, C-reactive protein.")}
if (nrow(crp_endpoint) > 0) {
  crp_endpoint %>%
    mutate(study_id = reorder(factor(as.character(study_id)), -max_pct_decline)) %>%
    ggplot(aes(x = study_id, y = max_pct_decline)) +
    geom_col(width = 0.75, fill = "#2C6FAC") +
    geom_hline(yintercept = 0, linewidth = 0.4) +
    scale_y_continuous(expand = expansion(mult = c(0, 0.08))) +
    labs(
      x = "Study ID",
      y = "Maximum CRP decline (%)"
    ) +
    theme_minimal(base_size = 12) +
    theme(panel.grid.major.x = element_blank())
} else {
  plot.new()
  text(0.5, 0.5, "No evaluable CRP endpoint data")
}
```

```{r tumor-waterfall-plot, fig.width = 10, fig.height = 5.5, fig.cap=fig_cap("Best percent change in target-lesion sum of diameters (SOD) among patients with ≥2 on-study scans (n = 5). All patients were on the original continued-colchicine schedule (not the amended 2-week course). Companion annotation table follows. Abbreviations: SOD, sum of diameters.")}
# Prior IO: concurrent agents joined with "+"; sequential lines show latest only
io_rx_pat <- "(pembrolizumab|nivolumab|ipilimumab|atezolizumab|durvalumab|avelumab|cemiplimab)"
io_display_order <- c(
  "ipilimumab", "nivolumab", "pembrolizumab", "atezolizumab",
  "durvalumab", "avelumab", "cemiplimab"
)
normalize_io <- function(x) {
  dplyr::recode(
    tolower(x),
    pembrolizumab = "pembrolizumab",
    nivolumab = "nivolumab",
    ipilimumab = "ipilimumab",
    atezolizumab = "atezolizumab",
    .default = tolower(x)
  )
}
format_latest_io_regimen <- function(drugs, start_dt, end_dt) {
  drugs <- normalize_io(drugs)
  n <- length(drugs)
  if (n == 0) return(NA_character_)
  if (n == 1) return(drugs[[1]])
  ord <- order(start_dt, drugs)
  drugs <- drugs[ord]
  start_dt <- start_dt[ord]
  end_dt <- end_dt[ord]
  line_id <- integer(n)
  line_id[[1]] <- 1L
  cur_line <- 1L
  for (i in seq_len(n)[-1]) {
    overlaps_cur <- any(
      start_dt[[i]] <= end_dt[line_id == cur_line] &
        start_dt[line_id == cur_line] <= end_dt[[i]]
    )
    if (overlaps_cur) {
      line_id[[i]] <- cur_line
    } else {
      cur_line <- cur_line + 1L
      line_id[[i]] <- cur_line
    }
  }
  line_end <- vapply(unique(line_id), function(L) max(end_dt[line_id == L]), as.Date(NA))
  latest_line <- unique(line_id)[which.max(line_end)]
  latest_drugs <- unique(drugs[line_id == latest_line])
  latest_drugs <- latest_drugs[order(match(latest_drugs, io_display_order))]
  paste(latest_drugs, collapse = " + ")
}
# Calendar days from most recent cycle IP start (not a fixed-length protocol visit day)
cycle_day_label <- function(dt, cycle_df) {
  if (is.na(dt) || nrow(cycle_df) == 0) return(NA_character_)
  eligible <- cycle_df %>% dplyr::filter(.data$cycle_start <= dt)
  if (nrow(eligible) == 0) return(NA_character_)
  row <- eligible %>% dplyr::slice_max(.data$cycle_start, n = 1, with_ties = FALSE)
  day <- as.integer(dt - row$cycle_start[[1]]) + 1L
  paste0("C", row$cycle[[1]], "D", day)
}

cycle_starts <- dat %>%
  filter(
    study_id %in% as.character(tumor_change$study_id),
    !is.na(date_ip_dispensed)
  ) %>%
  mutate(
    study_id = as.character(study_id),
    cycle = dplyr::case_when(
      stringr::str_detect(redcap_event_name, "cycle_1") ~ 1L,
      stringr::str_detect(redcap_event_name, "cycle_2") ~ 2L,
      stringr::str_detect(redcap_event_name, "cycle_3") ~ 3L,
      stringr::str_detect(redcap_event_name, "cycle_4") ~ 4L,
      stringr::str_detect(redcap_event_name, "cycle_5") ~ 5L,
      stringr::str_detect(redcap_event_name, "cycle_6") ~ 6L,
      TRUE ~ NA_integer_
    )
  ) %>%
  filter(!is.na(cycle)) %>%
  group_by(study_id, cycle) %>%
  summarise(cycle_start = min(date_ip_dispensed), .groups = "drop")

tx_index <- dat %>%
  filter(study_id %in% as.character(tumor_change$study_id)) %>%
  group_by(study_id = as.character(study_id)) %>%
  summarise(
    index_date = dplyr::coalesce(
      suppressWarnings(min(.data$date_ip_dispensed, na.rm = TRUE)),
      suppressWarnings(min(.data$date_of_enrollment, na.rm = TRUE))
    ),
    .groups = "drop"
  ) %>%
  mutate(index_date = if_else(is.infinite(.data$index_date), as.Date(NA), .data$index_date))

prior_io <- dat %>%
  filter(
    study_id %in% as.character(tumor_change$study_id),
    redcap_repeat_instrument == "prior_treatment_therapy",
    !is.na(prior_treatment_therapies),
    prior_treatment_therapies != ""
  ) %>%
  mutate(
    study_id = as.character(study_id),
    therapy_l = tolower(coalesce(prior_treatment_therapies, "")),
    drug = stringr::str_extract(.data$therapy_l, io_rx_pat),
    start_dt = .data$therapy_start_date,
    end_dt = dplyr::coalesce(.data$therapy_end_date, .data$therapy_start_date)
  ) %>%
  filter(!is.na(.data$drug)) %>%
  arrange(study_id, .data$start_dt, .data$drug) %>%
  group_by(study_id) %>%
  summarise(
    io_regimen = format_latest_io_regimen(.data$drug, .data$start_dt, .data$end_dt),
    last_io_dose = max(.data$end_dt, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  left_join(tx_index, by = "study_id") %>%
  mutate(
    months_since_last_io = as.numeric(.data$index_date - .data$last_io_dose) / 30.44
  )

scan_cd_map <- tumor_change %>%
  mutate(study_id = as.character(study_id)) %>%
  left_join(
    dat %>%
      filter(study_id %in% as.integer(tumor_change$study_id)) %>%
      transmute(
        study_id = as.character(study_id),
        assess_date = dplyr::coalesce(
          .data$date_of_target_lesions_eva,
          .data$date_of_overall_response
        )
      ) %>%
      filter(!is.na(.data$assess_date)) %>%
      distinct() %>%
      arrange(.data$study_id, .data$assess_date) %>%
      group_by(.data$study_id) %>%
      summarise(first_scan_date = dplyr::first(.data$assess_date), .groups = "drop"),
    by = "study_id"
  ) %>%
  rowwise() %>%
  mutate(
    first_scan_cd = cycle_day_label(
      first_scan_date,
      cycle_starts %>% filter(study_id == .env$study_id)
    ),
    scan_cd = cycle_day_label(
      imaging_date,
      cycle_starts %>% filter(study_id == .env$study_id)
    ),
    crp_cd = cycle_day_label(
      crp_date_at_imaging,
      cycle_starts %>% filter(study_id == .env$study_id)
    ),
    best_is_first_scan = !is.na(first_scan_date) &
      !is.na(imaging_date) &
      first_scan_date == imaging_date
  ) %>%
  ungroup() %>%
  select(study_id, first_scan_cd, scan_cd, crp_cd, best_is_first_scan)

tumor_waterfall_dat <- tumor_change %>%
  mutate(study_id = as.character(study_id)) %>%
  left_join(prior_io, by = "study_id") %>%
  left_join(scan_cd_map, by = "study_id") %>%
  mutate(
    plot_y = best_pct_change,
    direction = if_else(best_pct_change < 0, "Shrinkage", "Growth"),
    bar_lab = sprintf("%+.0f%%", best_pct_change),
    crp_delta_lab = case_when(
      is.na(crp_pct_change_at_imaging) ~ "NE",
      crp_pct_change_at_imaging >= 0 ~
        paste0("↓", sprintf("%.0f", crp_pct_change_at_imaging), "%"),
      TRUE ~
        paste0("↑", sprintf("%.0f", abs(crp_pct_change_at_imaging)), "%")
    ),
    prior_io_lab = if_else(
      is.na(io_regimen) | io_regimen == "",
      "—",
      paste0(io_regimen, " (", sprintf("%.1f", months_since_last_io), " mo)")
    )
  ) %>%
  mutate(study_id = reorder(factor(study_id), best_pct_change))

if (nrow(tumor_waterfall_dat) > 0) {
  p_tumor_waterfall <- ggplot(
    tumor_waterfall_dat,
    aes(x = study_id, y = plot_y, fill = direction)
  ) +
    geom_col(width = 0.7) +
    geom_hline(yintercept = 0, linewidth = 0.5) +
    geom_text(
      aes(
        label = bar_lab,
        vjust = if_else(plot_y >= 0, -0.35, 1.35)
      ),
      size = 5,
      fontface = "bold"
    ) +
    scale_fill_manual(
      values = c("Shrinkage" = "#2C6FAC", "Growth" = "#C75B39"),
      guide = "none"
    ) +
    scale_y_continuous(expand = expansion(mult = c(0.12, 0.14))) +
    labs(
      title = "Best % change in target-lesion SOD (n = 5)",
      subtitle = "Original continued colchicine schedule (not amended 2-week dosing)",
      x = "Study ID",
      y = "Best % change from baseline",
      caption = "Annotation details (scan timing, CRP, prior IO) are in the table below."
    ) +
    theme_minimal(base_size = 18) +
    theme(
      panel.grid.major.x = element_blank(),
      plot.title = element_text(size = 20, face = "bold", hjust = 0),
      plot.subtitle = element_text(size = 14, face = "bold", color = "#1F4E79", hjust = 0),
      plot.caption.position = "plot",
      plot.caption = element_text(size = 12, hjust = 0, color = "#333333", margin = margin(t = 10)),
      axis.title = element_text(size = 16, face = "bold"),
      axis.text = element_text(size = 15),
      plot.margin = margin(10, 14, 12, 10)
    )
  print(p_tumor_waterfall)
} else {
  p_tumor_waterfall <- NULL
  plot.new()
  text(0.5, 0.5, "No evaluable tumor-change data")
}
```

```{r tumor-waterfall-table}
tumor_annot_tbl <- tumor_waterfall_dat %>%
  arrange(best_pct_change) %>%
  transmute(
    `Study ID` = as.character(study_id),
    `Best SOD %` = sprintf("%+.1f", best_pct_change),
    `First scan` = dplyr::coalesce(first_scan_cd, "NE"),
    `Best scan` = dplyr::coalesce(scan_cd, "NE"),
    `CRP baseline` = if_else(is.na(baseline_crp), "NE", sprintf("%.1f", baseline_crp)),
    `CRP at best-scan draw` = if_else(is.na(crp_at_imaging), "NE", sprintf("%.1f", crp_at_imaging)),
    `ΔCRP %` = crp_delta_lab,
    `CRP draw timing` = dplyr::coalesce(crp_cd, "NE"),
    `Prior IO (mo before start)` = prior_io_lab
  )

tumor_annot_tbl %>%
  kbl(caption = tbl_cap(paste0(
    "Annotation for Figure 2 (waterfall). All patients on original continued-dose schedule. ",
    "C#D# = calendar days from that cycle’s IP start (may extend past the 14-day dosing block into follow-up). ",
    "CRP at best-scan draw = CRP nearest the best SOD imaging date (±21 days). ",
    "Prior IO: concurrent agents joined with “+”; sequential lines show latest agent only. ",
    "Abbreviations: SOD, sum of diameters; CRP, C-reactive protein (mg/L); IO, immunotherapy; NE, not evaluable."
  ))) %>%
  ctr_table_style() %>%
  kableExtra::footnote(
    general = "Rows ordered left-to-right as in the waterfall (best to worst SOD change).",
    general_title = ""
  )
```

```{r crp-longitudinal-plot, fig.width = 9, fig.height = 11, fig.cap = fig_cap("Longitudinal peripheral blood CRP (mg/L) across protocol C1 visits (baseline C1D1 and post-baseline C1D8/C1D15) for enrolled patients with serial C1 values. Abbreviations: CRP, C-reactive protein.")}
crp_plot <- crp_long %>%
  mutate(
    visit = case_when(
      redcap_event_name %in% c("cycle_1_day_1__3_d_arm_1", "cycle_1_day_1_arm_2", "cycle_1_day_1_arm_3") ~ "C1D1",
      redcap_event_name %in% c("cycle_1_day_8_arm_2", "cycle_1_day_8_arm_3") ~ "C1D8",
      redcap_event_name %in% c("cycle_1_day_15__3_arm_1", "cycle_1_day_15_arm_2", "cycle_1_day_15_arm_3") ~ "C1D15",
      TRUE ~ NA_character_
    ),
    study_id = factor(study_id),
    visit = factor(visit, levels = c("C1D1", "C1D8", "C1D15"))
  ) %>%
  filter(!is.na(visit)) %>%
  group_by(study_id, visit) %>%
  summarise(crp_value = median(crp_value, na.rm = TRUE), .groups = "drop") %>%
  group_by(study_id) %>%
  filter(n_distinct(visit) >= 2, any(visit %in% c("C1D8", "C1D15"))) %>%
  ungroup()

if (nrow(crp_plot) > 0) {
  ggplot(crp_plot, aes(x = visit, y = crp_value, group = study_id, color = study_id)) +
    geom_line(linewidth = 1.1, alpha = 0.85) +
    geom_point(size = 3.5) +
    scale_y_continuous(expand = expansion(mult = c(0.02, 0.06)), breaks = scales::pretty_breaks(n = 14)) +
    labs(x = "Visit", y = "CRP (mg/L)", color = "Study ID") +
    theme_minimal(base_size = 13) +
    theme(legend.position = "right")
} else {
  plot.new()
  text(0.5, 0.5, "No CRP longitudinal data")
}
```

---

## General Toxicity Profile

Treatment-related adverse events (TRAE; attribution coded as Possible/Probable/Definite) are summarized by **worst CTCAE grade per patient per preferred term** (if multiple grades are recorded for the same term, only the maximum grade is counted).

```{r toxicity-ae-patient-incidence, eval=nrow(ae_by_term_grade_patient) > 0}
ae_by_term_grade_patient %>%
  rename(`Adverse event (CTCAE v5)` = ae_term_display, Grade = ae_grade_label) %>%
  mutate(Grade = paste0("Grade ", Grade)) %>%
  select(`Adverse event (CTCAE v5)`, Grade, `N (%)`) %>%
  kbl(caption = tbl_cap(paste0(
    "Treatment-related adverse events by CTCAE v5 preferred term and worst grade per patient (N = ",
    enrolled_n,
    "). Patients with any grade \u22653 event: ",
    round(grade3plus_rate, 0),
    "%."
  ))) %>%
  ctr_table_style()
```

---

## Pharmacokinetics and Pharmacodynamics

Pharmacokinetics were not evaluated. Pharmacodynamic assessments included CRP (primary), ferritin, and erythrocyte sedimentation rate (ESR) when collected per protocol.

```{r pd-summary}
pd_long <- dat %>%
  filter(study_id %in% enrolled_ids) %>%
  select(study_id, redcap_event_name, crp_value, ferritin_value_num, esr_value_num) %>%
  pivot_longer(
    cols = c(crp_value, ferritin_value_num, esr_value_num),
    names_to = "marker",
    values_to = "value"
  ) %>%
  filter(!is.na(value)) %>%
  mutate(
    marker = recode(marker,
      crp_value = "CRP (mg/L)",
      ferritin_value_num = "Ferritin (µg/dL)",
      esr_value_num = "ESR (mm)"
    )
  )

pd_summary <- pd_long %>%
  group_by(marker) %>%
  summarise(
    N = n(),
    Median = median(value),
    IQR_low = quantile(value, 0.25),
    IQR_high = quantile(value, 0.75),
    .groups = "drop"
  ) %>%
  mutate(
    `Median (IQR)` = paste0(
      fmt_num(Median), " (", fmt_num(IQR_low), "–", fmt_num(IQR_high), ")"
    )
  ) %>%
  select(Marker = marker, N, `Median (IQR)`)

pd_summary %>%
  kbl(caption = tbl_cap("Pharmacodynamic laboratory values (all on-study assessments, enrolled patients).")) %>%
  ctr_table_style()
```

```{r pd-change-summary, results='asis'}
# PD change block: baseline (C1D1; screening fallback) to post-baseline C1 (C1D8/C1D15).
pd_marker_cols <- c(
  crp_value = "CRP (mg/L)",
  ferritin_value_num = "Ferritin (µg/dL)",
  esr_value_num = "ESR (mm)"
)

pd_base <- dat %>%
  filter(study_id %in% enrolled_ids, redcap_event_name %in% c(baseline_events, screening_events)) %>%
  select(study_id, redcap_event_name, all_of(names(pd_marker_cols))) %>%
  pivot_longer(
    cols = all_of(names(pd_marker_cols)),
    names_to = "marker_key",
    values_to = "value"
  ) %>%
  filter(!is.na(value)) %>%
  mutate(
    marker = recode(marker_key, !!!pd_marker_cols),
    source_priority = if_else(redcap_event_name %in% baseline_events, 0L, 1L)
  ) %>%
  arrange(study_id, marker, source_priority, redcap_event_name) %>%
  group_by(study_id, marker) %>%
  slice(1) %>%
  ungroup() %>%
  select(study_id, marker, baseline_value = value)

pd_post <- dat %>%
  filter(study_id %in% enrolled_ids, redcap_event_name %in% post_baseline_events) %>%
  select(study_id, redcap_event_name, all_of(names(pd_marker_cols))) %>%
  pivot_longer(
    cols = all_of(names(pd_marker_cols)),
    names_to = "marker_key",
    values_to = "value"
  ) %>%
  filter(!is.na(value)) %>%
  mutate(marker = recode(marker_key, !!!pd_marker_cols)) %>%
  group_by(study_id, marker) %>%
  summarise(
    post_min_value = min(value, na.rm = TRUE),
    n_post_values = n(),
    .groups = "drop"
  )

pd_change <- pd_base %>%
  inner_join(pd_post, by = c("study_id", "marker")) %>%
  filter(!is.na(baseline_value), baseline_value > 0) %>%
  mutate(
    pct_change = 100 * (post_min_value - baseline_value) / baseline_value,
    pct_decline = pmax(-pct_change, 0)
  )

pd_change_summary <- pd_change %>%
  group_by(marker) %>%
  summarise(
    `N evaluable` = n_distinct(study_id),
    `Baseline median` = fmt_num(median(baseline_value, na.rm = TRUE), 1),
    `Post-baseline minimum median` = fmt_num(median(post_min_value, na.rm = TRUE), 1),
    `Median % decline (IQR)` = paste0(
      fmt_num(median(pct_decline, na.rm = TRUE), 1), " (",
      fmt_num(quantile(pct_decline, 0.25, na.rm = TRUE), 1), "–",
      fmt_num(quantile(pct_decline, 0.75, na.rm = TRUE), 1), ")"
    ),
    .groups = "drop"
  ) %>%
  rename(Marker = marker)

pd_change_summary %>%
  kbl(caption = tbl_cap("Pharmacodynamic decline from baseline to minimum post-baseline C1 value (C1D8/C1D15). Decline is bounded at 0% (no decline).")) %>%
  ctr_table_style()

ferritin_c1_evaluable_n <- pd_change %>%
  filter(marker == "Ferritin (µg/dL)") %>%
  summarise(n = n_distinct(study_id)) %>%
  pull(n)

ferritin_any_post_evaluable_n <- {
  pd_post_any <- dat %>%
    filter(study_id %in% enrolled_ids) %>%
    filter(!redcap_event_name %in% c(screening_events, baseline_events)) %>%
    select(study_id, redcap_event_name, ferritin_value_num) %>%
    filter(!is.na(ferritin_value_num)) %>%
    distinct(study_id)

  pd_base_ferritin <- pd_base %>%
    filter(marker == "Ferritin (µg/dL)") %>%
    distinct(study_id)

  inner_join(pd_base_ferritin, pd_post_any, by = "study_id") %>%
    summarise(n = n_distinct(study_id)) %>%
    pull(n)
}

cat(
  "\n\n**Ferritin evaluable count note:** baseline-to-post C1 (D8/D15) ferritin was available for ",
  ferritin_c1_evaluable_n,
  " patient(s); using any post-baseline visit (any cycle) would yield ",
  ferritin_any_post_evaluable_n,
  " patient(s).\n\n",
  sep = ""
)
```

```{r pd-crp-responders}
crp_response <- pd_change %>%
  filter(marker == "CRP (mg/L)") %>%
  summarise(
    `N evaluable` = n(),
    `Decline ≥30%, n (%)` = paste0(sum(pct_decline >= 30), " (", sprintf("%.1f", 100 * mean(pct_decline >= 30)), "%)"),
    `Decline ≥50%, n (%)` = paste0(sum(pct_decline >= 50), " (", sprintf("%.1f", 100 * mean(pct_decline >= 50)), "%)"),
    `Median CRP % decline (IQR)` = paste0(
      fmt_num(median(pct_decline, na.rm = TRUE), 1), " (",
      fmt_num(quantile(pct_decline, 0.25, na.rm = TRUE), 1), "–",
      fmt_num(quantile(pct_decline, 0.75, na.rm = TRUE), 1), ")"
    )
  )

crp_response %>%
  kbl(caption = tbl_cap("CRP pharmacodynamic response summary among evaluable patients.")) %>%
  ctr_table_style()
```

---

## Patient Status and Survival

Progression-free survival (PFS) and overall survival (OS) were **not captured on the dedicated REDCap PFS/OS form** in this export (all `subject_progression_status` fields blank). Endpoints below were derived per patient from: (1) RECIST 1.1 best overall response (progressive disease [PD] date); (2) protocol therapy discontinuation for documented progression; (3) final subject status and date of death or last contact. Time zero = first investigational product dispense date, or enrollment if missing.

**Analyses in this section include only the `r km_eligible_n` patients enrolled under the original indefinite colchicine schedule with serial imaging surveillance (REDCap Arm 1).** The `r enrolled_amended_n` patients enrolled after the protocol amendment (fixed 2-week intervention; REDCap Arm 2) lacked the same longitudinal imaging framework and are reported for CRP pharmacodynamics and safety only.

```{r bor-summary}
bor_tbl <- surv_dat %>%
  transmute(`Best overall response (RECIST 1.1)` = coalesce(best_recist, "Not assessed")) %>%
  count(`Best overall response (RECIST 1.1)`, name = "N patients") %>%
  mutate(`N (%)` = paste0(`N patients`, " (", sprintf("%.1f", 100 * `N patients` / km_eligible_n), "%)")) %>%
  arrange(desc(`N patients`), `Best overall response (RECIST 1.1)`)

bor_tbl %>%
  kbl(caption = tbl_cap("Exploratory best overall response (RECIST 1.1) among original-schedule enrolled patients (Kaplan–Meier cohort).")) %>%
  ctr_table_style()
```

```{r pfs-os-summary}
surv_summary <- tibble(
  Endpoint = c("Progression-free survival", "Overall survival"),
  `N patients` = c(nrow(surv_pfs), nrow(surv_os)),
  `N events` = c(sum(surv_pfs$pfs_event), sum(surv_os$os_event)),
  `Median (months, 95% CI)` = c(
    if (is.na(pfs_med$median)) {
      "Not reached / not estimable"
    } else {
      paste0(
        fmt_num(pfs_med$median, 1), " (",
        fmt_num(pfs_med$lcl, 1), "–", fmt_num(pfs_med$ucl, 1), ")"
      )
    },
    if (is.na(os_med$median)) {
      "Not reached / not estimable"
    } else {
      paste0(
        fmt_num(os_med$median, 1), " (",
        fmt_num(os_med$lcl, 1), "–", fmt_num(os_med$ucl, 1), ")"
      )
    }
  )
)
surv_summary %>%
  kbl(caption = tbl_cap(paste0(
    "Exploratory progression-free survival and overall survival summary (Kaplan–Meier medians; original-schedule cohort, n = ",
    km_eligible_n, ")."
  ))) %>%
  ctr_table_style()
```

```{r pfs-os-by-patient}
pfs_os_tbl %>%
  kbl(
    caption = tbl_cap("Patient-level progression-free survival and overall survival among original-schedule enrolled patients (derived from RECIST, discontinuation, and final status fields).")
  ) %>%
  ctr_table_style()
```

```{r km-pfs, fig.cap=fig_cap(paste0("Progression-free survival (Kaplan–Meier) from treatment start among patients enrolled under the original indefinite colchicine schedule with surveillance imaging (n = ", km_eligible_n, "; exploratory; not powered). PFS events: RECIST PD or discontinuation for progression. Abbreviations: PFS, progression-free survival; PD, progressive disease; RECIST, Response Evaluation Criteria in Solid Tumors.")), eval=!is.null(fit_pfs)}
if (!is.null(fit_pfs)) {
  ggsurvplot(
    fit_pfs,
    data = surv_pfs,
    risk.table = TRUE,
    conf.int = TRUE,
    xlab = "Months from treatment start",
    ylab = "Progression-free survival probability",
    title = "Exploratory progression-free survival",
    legend = "none",
    risk.table.height = 0.25
  )
}
```

```{r km-os, fig.cap=fig_cap(paste0("Overall survival (Kaplan–Meier) from treatment start among patients enrolled under the original indefinite colchicine schedule with surveillance imaging (n = ", km_eligible_n, "; exploratory; not powered). Abbreviations: OS, overall survival.")), eval=!is.null(fit_os)}
if (!is.null(fit_os)) {
  ggsurvplot(
    fit_os,
    data = surv_os,
    risk.table = TRUE,
    conf.int = TRUE,
    xlab = "Months from treatment start",
    ylab = "Overall survival probability",
    title = "Exploratory overall survival",
    legend = "none",
    risk.table.height = 0.25
  )
}
```

*These survival analyses are descriptive and underpowered (Kaplan–Meier cohort n = `r km_eligible_n`; total enrolled n = `r enrolled_n`). Medians and curves should be interpreted cautiously; several patients lack formal final-status forms but have tumor assessment or visit dates used for censoring. Patients enrolled after the protocol amendment (n = `r enrolled_amended_n`) are excluded because they received a fixed 2-week intervention without the same longitudinal imaging surveillance framework.*

---

## Statistical Analysis

Analyses followed protocol v3.0 (Section 12). The primary endpoint was tested with a one-sided one-sample *t*-test (H~0~: mean maximum CRP decline = 0) at α = 0.025 per the Bonferroni-adjusted Cohort 1 low-dose analysis plan. Exploratory progression-free and overall survival estimates used the Kaplan–Meier method (`survival` package in R) among the `r km_eligible_n` patients enrolled under the original indefinite-treatment schedule with serial imaging surveillance (pre-amendment enrollments only). PFS and OS were derived from RECIST 1.1 assessments, protocol therapy discontinuation for progression, and final subject status because dedicated PFS/OS case report forms were incomplete in this export. Treatment-related adverse events required Possible, Probable, or Definite attribution to colchicine and were summarized by CTCAE v5.0 preferred term and worst grade per patient. Primary and exploratory analyses were performed in R per protocol v3.0 Section 12. Annotated R code and primary output are provided in the [statistical analysis file](https://alexanderb95-jpg.github.io/colchicine-ctr-stats/statistical-analysis.html).

---

## Discussion

This single-center pilot (STUDY-21-001175) evaluated whether short-course low-dose colchicine could suppress peripheral blood CRP in patients with advanced or recurrent solid tumors. The trial closed under planned accrual: Cohort 1 low dose enrolled `r enrolled_n` patients, the high-dose sequential arm never opened, and Cohort 2 (resected high-risk urothelial carcinoma planned for adjuvant nivolumab) closed before enrollment because of poor accrual.

Among `r consented_n` consenting patients, `r screen_fail_n` (`r fmt_num(100 * screen_fail_n / consented_n, 1)`%) were screen failures before enrollment. `r evaluable_efficacy_n` of `r enrolled_n` enrolled patients were evaluable for the primary endpoint. The mean maximum CRP decline was `r fmt_num(primary_mean)`% (SD `r fmt_num(primary_sd)`%). The prespecified one-sided one-sample *t*-test yielded *p* = `r fmt_p(primary_p)`, which does not meet the protocol’s multiplicity-adjusted α = 0.025 threshold.

With `r evaluable_efficacy_n` evaluable patients instead of the planned 15 per low-dose group, post-hoc power to detect the observed effect (Cohen’s d ≈ `r fmt_num(cohens_d_primary, 2)`) at α = 0.025 was approximately `r fmt_num(posthoc_power_pct, 0)`%.

Colchicine 0.6 mg twice daily for 14 days was well tolerated with grade 1–2 treatment-related adverse events only. Exploratory RECIST and Kaplan–Meier analyses were underpowered. These findings provide preliminary biologic proof-of-concept for inflammatory modulation but do not support colchicine as an effective re-sensitization strategy.

---

## Acknowledgments

[To be completed by investigators.]

## Funding

[To be completed.]

## Conflict of Interest

[To be completed.]

## Data Availability

De-identified data may be made available upon reasonable request to the corresponding author and subject to institutional approval.

## Author Contributions

[To be completed.]

## References

1. Doroshow DB, Galsky MD, Marron TU, et al. Protocol STUDY-21-001175, version 3.0 (22 May 2023). Icahn School of Medicine at Mount Sinai.  
2. Tran M, et al. [Pro-tumorigenic inflammation / urothelial carcinoma translational study; PMC13069984](https://pmc.ncbi.nlm.nih.gov/articles/PMC13069984/).  
3. Ridker PM, et al. CANTOS trial. *Lancet*. 2017.  
4. CheckMate 274: adjuvant nivolumab in high-risk urothelial carcinoma.  

---

## Figures and Tables

All primary tables and figures are embedded in the sections above. Additional correlative analyses (ctDNA, cytokines, single-cell myeloid signatures) were prespecified but are not included in this CTR results report pending separate correlative manuscripts.

## Appendix: Included Subjects Dataset

```{r appendix-included-subjects}
included_subject_ids <- sort(unique(crp_endpoint$study_id))

included_crp_visits <- crp_long %>%
  filter(study_id %in% included_subject_ids) %>%
  mutate(
    visit = case_when(
      redcap_event_name %in% c("cycle_1_day_1__3_d_arm_1", "cycle_1_day_1_arm_2", "cycle_1_day_1_arm_3") ~ "C1D1",
      redcap_event_name %in% c("cycle_1_day_8_arm_2", "cycle_1_day_8_arm_3") ~ "C1D8",
      redcap_event_name %in% c("cycle_1_day_15__3_arm_1", "cycle_1_day_15_arm_2", "cycle_1_day_15_arm_3") ~ "C1D15",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(visit)) %>%
  group_by(study_id, visit) %>%
  summarise(crp_mg_l = min(crp_value, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = visit, values_from = crp_mg_l, names_prefix = "CRP_")

included_tbl <- tibble(study_id = included_subject_ids) %>%
  left_join(
    patient_master %>%
      select(
        study_id,
        primary_tumor_type,
        age_years,
        gender_label,
        ecog_label
      ),
    by = "study_id"
  ) %>%
  left_join(prior_systemic_lines, by = "study_id") %>%
  left_join(
    crp_endpoint %>%
      select(study_id, baseline_crp_mg_l, min_on_tx_crp_mg_l, max_pct_decline, n_crp_on_tx),
    by = "study_id"
  ) %>%
  left_join(included_crp_visits, by = "study_id") %>%
  left_join(
    surv_dat %>%
      select(study_id, pfs_event, pfs_months, os_event, os_months),
    by = "study_id"
  ) %>%
  mutate(
    `Study ID` = study_id,
    `Primary cancer type` = primary_tumor_type,
    `Age (years)` = round(age_years, 0),
    Sex = gender_label,
    ECOG = ecog_label,
    `Prior systemic lines` = n_systemic_lines,
    `CRP baseline C1D1 (mg/L)` = round(baseline_crp_mg_l, 1),
    `CRP C1D8 (mg/L)` = round(CRP_C1D8, 1),
    `CRP C1D15 (mg/L)` = round(CRP_C1D15, 1),
    `Lowest post-baseline CRP (mg/L)` = round(min_on_tx_crp_mg_l, 1),
    `Max CRP decline (%)` = round(max_pct_decline, 1),
    `N post-baseline CRP values` = n_crp_on_tx,
    `PFS event` = if_else(pfs_event, "Yes", "No"),
    `PFS (months)` = round(pfs_months, 1),
    `OS event (death)` = if_else(os_event, "Yes", "No"),
    `OS (months)` = round(os_months, 1)
  ) %>%
  select(
    `Study ID`,
    `Primary cancer type`,
    `Age (years)`,
    Sex,
    ECOG,
    `Prior systemic lines`,
    `CRP baseline C1D1 (mg/L)`,
    `CRP C1D8 (mg/L)`,
    `CRP C1D15 (mg/L)`,
    `Lowest post-baseline CRP (mg/L)`,
    `Max CRP decline (%)`,
    `N post-baseline CRP values`,
    `PFS event`,
    `PFS (months)`,
    `OS event (death)`,
    `OS (months)`
  ) %>%
  arrange(`Study ID`)

included_tbl %>%
  kbl(
    caption = tbl_cap(paste0(
      "Subject-level dataset for patients included in primary endpoint analysis (N = ",
      nrow(included_tbl),
      ")."
    ))
  ) %>%
  ctr_table_style()
```

---

*Internal analytic report generated from REDCap export. Verify all numbers against source data and PI review before journal submission.*
